---
title: "Reproducibility Report: Group B Choice 2"
output:
  html_document:
    toc: true
    toc_float: true
---

-------

For this exercise, please try to reproduce the results from Experiment 1 of the associated paper (Ko, Sadler & Galinsky, 2015). The PDF of the paper is included in the same folder as this Rmd file. 

#### Methods summary: 

A sense of power has often been tied to how we perceive each other's voice. Social hierarchy is embedded into the structure of society and provides a metric by which others relate to one another. In 1956, the Brunswik Lens Model was introduced to examine how vocal cues might influence hierarchy. In "The Sound of Power: Conveying and Detecting Hierarchical Rank Through Voice," Ko and colleagues investigated how manipulation of hierarchal rank within a situation might impact vocal acoustic cues. Using the Brunswik Model, six acoustic metrics were utilized (pitch mean & variability, loudness mean & variability, and resonance mean & variability) to isolate a potential contribution between individuals of different hierarchal rank. In the first experiment, Ko, Sadler & Galinsky examined the vocal acoustic cues of individuals before and after being assigned a hierarchal rank in a sample of 161 subjects (80 male). Each of the six hierarchy acoustic cues were analyzed with a 2 (high vs. low rank condition) x 2 (male vs. female) analysis of covariance, controlling for the baseline of the respective acoustic cue.

------

#### Target outcomes: 

Below is the specific result you will attempt to reproduce (quoted directly from the results section of Experiment 1):

> The impact of hierarchical rank on speakers’ acoustic cues. Each of the six hierarchy-based (i.e., postmanipulation) acoustic variables was submitted to a 2 (condition: high rank, low rank) × 2 (speaker’s sex: female, male) between-subjects analysis of covariance, controlling for the corresponding baseline acoustic variable. Table 4 presents the adjusted means by condition. Condition had a significant effect on pitch, pitch variability, and loudness variability. Speakers’ voices in the high-rank condition had higher pitch, **F(1, 156) = 4.48, p < .05**; were more variable in loudness, **F(1, 156) = 4.66, p < .05**; and were more monotone (i.e., less variable in pitch), **F(1, 156) = 4.73, p < .05**, compared with speakers’ voices in the low-rank condition (**all other Fs < 1**; see the Supplemental Material for additional analyses of covariance involving pitch and loudness). (from Ko et al., 2015, p. 6; emphasis added)

The adjusted means for these analyses are reported in Table 4 (Table4_AdjustedMeans.png, included in the same folder as this Rmd file).

------

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Step 1: Load packages

```{r}
library(tidyverse) # for data munging
library(knitr) # for kable table formating
library(haven) # import and export 'SPSS', 'Stata' and 'SAS' Files
library(readxl) # import excel files

# #optional packages:
library(psych)
library(car) # for ANCOVA
library(compute.es) # for ANCOVA
library(lsmeans) # for ANCOVA

library(rstatix)
library(broom)
```

## Step 2: Load data

```{r}
# Just Experiment 1
d <-read_csv("data/S1_voice_level_Final.csv")
# DT::datatable(d)
```

## Step 3: Tidy data

```{r}
d_tidy = subset(d, select = -c(age, native, sex, voice, pow, feelpower, race, pitch_rmeanMD, pitch_rvarMD, intense_rmeanMD, intense_rvarMD, formant_rmeanMD, formant_rvarMD, pitch_smeanMD, pitch_svarMD, intense_smeanMD, intense_svarMD, formant_smeanMD, formant_svarMD, Zpitch_rmean, Zpitch_rvar, Zform_rmean, Zform_rvar, Zintense_rmean, Zintense_rvar, Zpitch_smean, Zpitch_svar, Zform_smean, Zform_svar, Zintense_smean, Zintense_svar))
d_tidy
```

## Step 4: Run analysis

### Pre-processing

```{r}
d_tidy = d_tidy %>%
 mutate(
  pitch_mean_adj = pitch_smean - pitch_rmean,
  pitch_var_adj = pitch_svar - pitch_rvar,
  intense_mean_adj = intense_smean - intense_rmean,
  intense_var_adj = intense_svar - intense_rvar,
  form_mean_adj = form_smean - form_rmean,
  form_var_adj = form_svar - form_rvar,
  )# calculate difference between experimental script and baseline for each voice
```

### Descriptive statistics

In the paper, the adjusted means by condition are reported (see Table 4, or Table4_AdjustedMeans.png, included in the same folder as this Rmd file). Reproduce these values below:

```{r}
d_tidy = d_tidy %>%
  group_by(plev) %>%
  mutate(
    avg_pitch_mean = mean(pitch_mean_adj) + mean(pitch_rmean),
    avg_pitch_var = mean(pitch_var_adj) + mean(pitch_rvar),
    avg_form_mean = mean(form_mean_adj) + mean(form_rmean),
    avg_form_var = mean(form_var_adj) + mean(form_rvar),
    avg_intense_mean = mean(intense_mean_adj) + mean(intense_rmean),
    avg_intense_var = mean(intense_var_adj) + mean(intense_rvar)
  ) %>% # average differences for each acoustic cue and then add to average baseline value
  ungroup()
d_tidy
```

### Inferential statistics

> The impact of hierarchical rank on speakers’ acoustic cues. Each of the six hierarchy-based (i.e., postmanipulation) acoustic variables was submitted to a 2 (condition: high rank, low rank) × 2 (speaker’s sex: female, male) between-subjects analysis of covariance, controlling for the corresponding baseline acoustic variable. [...] Condition had a significant effect on pitch, pitch variability, and loudness variability. Speakers’ voices in the high-rank condition had higher pitch, **F(1, 156) = 4.48, p < .05**; were more variable in loudness, **F(1, 156) = 4.66, p < .05**; and were more monotone (i.e., less variable in pitch), **F(1, 156) = 4.73, p < .05**, compared with speakers’ voices in the low-rank condition (**all other Fs < 1**; see the Supplemental Material for additional analyses of covariance involving pitch and loudness).

```{r}
# reproduce the above results here

#pitch_mean_ancova <- d_tidy %>%
#  anova_test(avg_pitch_mean ~ vsex + plev)
#get_anova_table(pitch_mean_ancova)

#pitch_var_ancova <- d_tidy %>%
#  anova_test(avg_pitch_var ~ vsex + plev)
#get_anova_table(pitch_var_ancova)

#intense_var_ancova <- d_tidy %>%
#  anova_test(avg_intense_var ~ vsex + plev)
#get_anova_table(intense_var_ancova)

# stopped here because reached 3 hours mark
```

# Step 5: Reflection

Were you able to reproduce the results you attempted to reproduce? If not, what part(s) were you unable to reproduce?

> I was able to only partially reproduce the results. For Table 4, the adjusted means I calculated were slightly off, and I'm not sure why. I tried different ways of adjusting the means based on the baseline acoustic cues, but I could not reproduce the exact values from the paper because the authors did not describe exactly how they adjusted the means. For the ANCOVA tests, I ran into issues with code syntax and was not able to get the test to run.

How difficult was it to reproduce your results? 

> Difficult, I was not able to reproduce all the results.

What aspects made it difficult? What aspects made it easy?

> The authors did not include specific details about how they calculated their adjustments for the means based on the baseline acoustic cues. I appreciated that the authors included a codebook document; it made it much easier to understand the variables (the Group A article I reproduced did not have a codebook). 